Homicide Increases in the COVID-19 Era

### 2025-02-20 Draft1.qmd
Added Plot 5 first draft

# Secondary datasets
for (i in 2018:2023) {
  obj_name <- paste0("california_median_income_", i)
  assign(obj_name, read.csv(paste0("Raw_Data/ACSST5Y",i,".S1903-Data.csv")))
}
```

@@

crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)

crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
                                  format = "%m/%d/%Y %I:%M:%S %p", 
                                  tz = "America/Los_Angeles")

@@

>>>>

# Combine the two datasets
crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)

# Secondary datasets
california_median_income_2018_2023 <- list()
yrs <- 2018:2023
for (i in 1:length(yrs)) {
  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
  obj <- obj[-1,]
  obj$year <- yrs[[i]]
  california_median_income_2018_2023[[i]] <- obj
}

cali_median_income <- do.call(rbind, california_median_income_2018_2023)

# adding a variable that's just the ZCTA zip code
cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
```

@@

```{r}
crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
                                  format = "%m/%d/%Y %I:%M:%S %p", 
                                  tz = "America/Los_Angeles")

@@

```{r}
library(sf)
library(tidygeocoder)
zcta <- read_sf(dsn = "cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

colnames(LA_zctas) <- c("X", "ZCTA5CE20")
LA_zctas$ZCTA5CE20 <- as.character(LA_zctas$ZCTA5CE20)

zcta_filtered <- zcta %>% inner_join(LA_zctas, by = "ZCTA5CE20")

median_household_income <- cali_median_income %>% select(ZCTA, S1903_C03_001E, year)

median_household_income$S1903_C03_001E <- as.integer(median_household_income$S1903_C03_001E)

zcta_with_data <- zcta_filtered %>% inner_join(median_household_income, join_by("ZCTA5CE20" == "ZCTA"))

plot(st_geometry(zcta_with_data), 
     col = brewer.pal(9, "Blues")[as.numeric(cut(zcta_with_data$S1903_C03_001E, 
                                                 breaks = 9))], 
     border = 'grey', 
     axes = TRUE, 
     main = "ZCTA Data Colored by Numerical Column", 
     cex.main = 1.5)
```
